Paper figures

### 2019-01-06 paper_figures.Rmd
added panel E

## Panel C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.
```{r DESeq2 loading, cache=TRUE}
pds_genus <- phyloseq_to_deseq2(gg_topi.genus, design = ~timepoint)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE)
res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
res.filt_genus <- res_genus[!is.na(res_genus$padj),]
res.filt_genus <- res.filt_genus[res.filt_genus$padj<0.05,]

@@

## Panel D
```{r}

>>>>

## Panel C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.
```{r DESeq2 loading, cache=TRUE}
pds_genus <- phyloseq_to_deseq2(gg_topi.genus, design = ~timepoint)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE)
```

```{r DESeq2 T2 T3}
res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
res.filt_genus <- res_genus[!is.na(res_genus$padj),]
res.filt_genus <- res.filt_genus[res.filt_genus$padj<0.05,]

@@

## Panel D: Ordination centroids beta diversity all timepoints 
```{r}
x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)

ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 0.8) +
  geom_point(data=centroids, size = 4) + 
  theme_minimal()
```

## Panel ?: Ordination beta diversity T3 vs T7
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
```

```{r, cache=TRUE}
ggtopi_subs <-subset_samples(gg_topi, timepoint %in% c('T3','T7'))
ggtopi_subs.relab <- transform_sample_counts(ggtopi_subs, function(x) x/sum(x))
ggtopi_subs.genus <- tax_glom(ggtopi_subs,taxrank = "Genus")
ggtopi_subs.genus.relab <- transform_sample_counts(ggtopi_subs.genus, function(x) x/sum(x))

mmtx <- model.matrix(object = ~timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
idx <- which(apply(mmtx, 2, function(x) all(x==0)))
mmtx <- mmtx[,-idx]

pds_37 <- phyloseq_to_deseq2(ggtopi_subs, design = ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_37 <- DESeq2::DESeq(pds_37, parallel = TRUE, quiet = TRUE, full = mmtx)

pds_genus <- phyloseq_to_deseq2(ggtopi_subs.genus, design =  ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE, full = mmtx)
```
## Panel E: Boxplot of abundance of significant genera from DESeq2 differential analysis
```{r}
res37 <- DESeq2::results(dds_genus, name = "timepointT7", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]

t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_subs), perl = TRUE)

genus_ab_t <- otu_table(ggtopi_subs.genus.relab)[rownames(res37.filt), ] %>%
            data.frame %>%
            tibble::rownames_to_column("OTU") %>%
            tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
            mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

genus_ab_t %>%

  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
ggplot() +
  geom_boxplot(aes(reorder(interaction(Family, Genus), Abundance, FUN = median), Abundance, fill=timepoint)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
  coord_flip() + 
  theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
      strip.text.y = element_text(angle = 0))
